---
title: "Data Prep - Fiji AIS"
author: "Avin Datt - PacifImpact"
date: "04 September 2020"
output: 
  github_document
---

# Scope 

AIS data was acquired via UNGP but due to the fact cluster resources were scarce, AIS data had to be extracted from Spark and then stored in a database for the wider PacifImpact team to access. 

# Load Libraries

```{r load libs}
# Load data.table library
if (!require(data.table)) {
  install.packages("data.table")
  library(data.table)
}

# Load DBI library
if (!require(DBI)) {
  install.packages("DBI")
  library(DBI)
}

# Load DBI library
if (!require(RPostgres)) {
  install.packages("RPostgres")
  library(RPostgres)
}

```

# Read AIS Suva

Read the files extracted from UNGP for Suva, Fiji Islands. 

```{r}
# Suva files
file.list <- list.files(path = "./data/external/ais_fj_suva/", 
                        pattern='*.csv')


# Read files in bulk
suva.list <- lapply(paste0("./data/external/ais_fj_suva/",
                           file.list),
                  fread)


# turn into one table 
suva.list <- rbindlist(suva.list)


head(suva.list)
```

# Write AIS Suva to PostgreSQL

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# # Append ws.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "stg_ais",
#                         suva.list,
#                         append = FALSE,
#                         row.names = FALSE)


rm(suva.list)

```

# Read AIS Lautoka

Read the files extracted from UNGP for Lautoka, Fiji Islands. 

```{r}
# Suva files
file.list <- list.files(path = "./data/external/ais_fj_lautoka/", 
                        pattern='*.csv')


# Read files in bulk
lautoka.list <- lapply(paste0("./data/external/ais_fj_lautoka/",
                           file.list),
                  fread)


# turn into one table 
lautoka.list <- rbindlist(lautoka.list)


head(lautoka.list)
```

# Write AIS Lautoka to PostgreSQL

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# # Append ws.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "stg_ais",
#                         lautoka.list,
#                         append = TRUE,
#                         row.names = FALSE)


count <- dbGetQuery(db.con, 'select count(*) from stg_ais')
# 8841715 records
```

 

