---
title: "Data Prep - Vanuatu AIS"
author: "Avin Datt - PacifImpact"
date: "04 September 2020"
output: 
  github_document
---


# Scope 

AIS data was acquired via UNGP but due to the fact cluster resources were scarce, AIS data had to be extracted from Spark and then stored in a database for the wider PacifImpact team to access. 

# Load Libraries

```{r load libs}
# Load data.table library
if (!require(data.table)) {
  install.packages("data.table")
  library(data.table)
}

# Load DBI library
if (!require(DBI)) {
  install.packages("DBI")
  library(DBI)
}

# Load DBI library
if (!require(RPostgres)) {
  install.packages("RPostgres")
  library(RPostgres)
}

```

# Read AIS Port Vila

Read the files extracted from UNGP for Port Vila, Vanuatu. 

```{r}
# port.v.list files
file.list <- list.files(path = "./data/external/ais_vn_port_vila/", 
                        pattern='*.csv')


# Read files in bulk
port.v.list <- lapply(paste0("./data/external/ais_vn_port_vila/",
                           file.list),
                  fread)


# turn into one table 
port.v.list <- rbindlist(port.v.list)


head(port.v.list)
```

# Write AIS Port Vila to PostgreSQL

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# # Append port.v.list to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "stg_ais",
#                         port.v.list,
#                         append = TRUE,
#                         row.names = FALSE)


rm(port.v.list)

```

# Read AIS Luganville

Read the files extracted from UNGP for Luganville, Vanuatu. 

```{r}
# lugv.list files
file.list <- list.files(path = "./data/external/ais_vn_luganville/", 
                        pattern='*.csv')


# Read files in bulk
lugv.list <- lapply(paste0("./data/external/ais_vn_luganville/",
                           file.list),
                  fread)


# turn into one table 
lugv.list <- rbindlist(lugv.list)


head(lugv.list)
```

# Write AIS Luganville to PostgreSQL

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Append lugv.list to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "stg_ais",
#                         lugv.list,
#                         append = TRUE,
#                         row.names = FALSE)


count <- dbGetQuery(db.con, 'select count(*) from stg_ais')
# 14013983 records
```

 

