---
title: "Data Prep - Pacific Islands Trade"
author: "Avin Datt - PacifImpact"
date: "03 September 2020"
output: 
  github_document
---

# Scope  

Ingest IMTS Data available from SPC Statistics for Development Division into PostgreSQL database for the following countries: 

- Fiji 
- Cook Islands 
- Solomon Islands
- Kiribati 
- Palau

IMTS Data for Vanuatu and Samoa available via the SPC Statistics for Development Division do not meet the projects eligibility criteria. As such, their appropriate trade data will be gathered from alternate sources, referenced below. 

# Reference

IMTS data for Fiji, Cook Islands and Solomon Islands is available in .xlsx format. It can be downloaded from [SPC Statistics for Development Division](https://sdd.spc.int/contact-us#disclaimer).

Eligible trade data for Samoa can be downloaded from [Samoa Bureau of Statistics](https://www.sbs.gov.ws/images/sbs-documents/Economics/Merchandise-Trade-Report/2020/OVERSEAS_MERCHANDISE_TRADE_JUNE_2020.pdf).

Eligible IMTS trade data for Vanuatu can be downloaded from [Vanuatu National Statistics Office](https://vnso.gov.vu/index.php/statistics-by-topic/trade#latest-trade-news).

# Load Libraries 

```{r load libs}
# Load openxlsx library 
if (!require(openxlsx)) {
  install.packages("openxlsx")
  library(openxlsx)
}

# Load data.table library
if (!require(data.table)) {
  install.packages("data.table")
  library(data.table)
}

# Load DBI library
if (!require(DBI)) {
  install.packages("DBI")
  library(DBI)
}

# Load DBI library
if (!require(RPostgres)) {
  install.packages("RPostgres")
  library(RPostgres)
}

```

# Read IMTS Balance of Trade Fiji Data 

Read in Balance of Trade - All Items for Fiji.

```{r}
# Read in Fiji File 
fj.imts.bot <- read.xlsx(xlsxFile = './data/external/Fiji_IMTS_Tables_2020.xlsx',
                         sheet = 1, 
                         startRow = 28,
                         skipEmptyCols = TRUE,
                         detectDates = TRUE,
                         rows = c(28:111),
                         cols = c(1:7))

head(fj.imts.bot)
```

Add and fix missing variable names. 

```{r}
# fj.imts.bot to data.table 
fj.imts.bot <- data.table(fj.imts.bot)


# Fix Monthly label to Year
setnames(fj.imts.bot, 
         "Monthly",
         "year")


# X2 should be Month 
setnames(fj.imts.bot, 
         "X2",
         "month")


# Exports FOB Domestic 
setnames(fj.imts.bot, 
         "X3",
         "exports-fob-domestic")


# Exports FOB Re-Export 
setnames(fj.imts.bot, 
         "X4",
         "exports-fob-reexport")


# Exports FOB Total 
setnames(fj.imts.bot, 
         "X5",
         "exports-fob-total")


# Imports CIF
setnames(fj.imts.bot, 
         "X6",
         "imports-cif")



# Trade Balance
setnames(fj.imts.bot, 
         "X7",
         "trade-balance")


head(fj.imts.bot)
```

Update Year values.

```{r}
# Update by using row number references
# 2014
fj.imts.bot[1:12, 
            year := 2014]


# 2015
fj.imts.bot[13:24, 
            year := 2015]


# 2016
fj.imts.bot[25:36, 
            year := 2016]


# 2017
fj.imts.bot[37:48, 
            year := 2017]


# 2018
fj.imts.bot[49:60, 
            year := 2018]


# 2019
fj.imts.bot[61:72, 
            year := 2019]


# 2020
fj.imts.bot[73:77, 
            year := 2020]


head(fj.imts.bot)
```

Melt the table to adhere to the key-value pair schema design. 

```{r}
# Melt the data to adhere to the schema design
fj.imts.bot <- melt.data.table(fj.imts.bot,
                               id = c("year",
                                      "month"), measure = c("exports-fob-domestic",
                                                            "exports-fob-reexport", 
                                                            "exports-fob-total",
                                                            "imports-cif",
                                                            "trade-balance"))


head(fj.imts.bot)
```

Create Date from year and month. 

```{r}
# Concatenate
fj.imts.bot[, 
            date := paste0(gsub(" ", "", year, fixed = TRUE),
                           "-",
                           gsub(" ", "",month, fixed = TRUE),
                           "-",
                           "01")]


# Format Date
fj.imts.bot[, 
            date := as.Date(date, format = "%Y-%B-%d")]


# Remove year and month fields 
fj.imts.bot[, 
            year := NULL][, 
                          month := NULL]


head(fj.imts.bot)
```

Rename "variable" to "name". 

```{r}
# Rename variable to name 
setnames(fj.imts.bot, 
         "variable",
         "name")


head(fj.imts.bot)
```

Add metric key, frequency, country, category, source and property informations. 

```{r}
# Metric 
fj.imts.bot[, 
            `:=` (frequency = "monthly", 
                  country = "fj", 
                  category = "trade", 
                  source = "imts", 
                  properties = "{'currency' = 'fjd 000'}",
                  metric_key = paste0("fj","-",name,"-",date))]


head(fj.imts.bot)
```

Reorder data. 

```{r}
# Change order of columns to align with schema
fj.imts.bot <- fj.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]


head(fj.imts.bot)
```

# Read IMTS Balance of Trade Cook Islands Data 

Read in Balance of Trade - All Items for Cook Islands.

```{r}
# Read in Fiji File 
ck.imts.bot <- read.xlsx(xlsxFile = './data/external/Cook_IMTS_Tables_2020.xlsx',
                         sheet = 1, 
                         startRow = 15,
                         skipEmptyCols = TRUE,
                         detectDates = TRUE,
                         rows = c(15:58),
                         cols = c(1:7))


head(ck.imts.bot)
```

Add and fix missing variable names. 

```{r}
# ck.imts.bot to data.table 
ck.imts.bot <- data.table(ck.imts.bot)


# Fix Monthly label to Year
setnames(ck.imts.bot, 
         "Monthly",
         "year")


# X2 should be Month 
setnames(ck.imts.bot, 
         "X2",
         "month")


# Exports FOB Domestic 
setnames(ck.imts.bot, 
         "X3",
         "exports-fob-domestic")


# Exports FOB Re-Export 
setnames(ck.imts.bot, 
         "X4",
         "exports-fob-reexport")


# Exports FOB Total 
setnames(ck.imts.bot, 
         "X5",
         "exports-fob-total")


# Imports CIF
setnames(ck.imts.bot, 
         "X6",
         "imports-cif")



# Trade Balance
setnames(ck.imts.bot, 
         "X7",
         "trade-balance")


head(ck.imts.bot)
```

Remove rogue empty row. 

```{r}
# Remove rows with no month value 
ck.imts.bot <- ck.imts.bot[!is.na(month)]


head(ck.imts.bot)
```


Update Year values.

```{r}
# Update by using row number references
# 2017
ck.imts.bot[1:12, 
            year := 2017]


# 2015
ck.imts.bot[13:24, 
            year := 2018]


# 2016
ck.imts.bot[25:36, 
            year := 2019]


# 2017
ck.imts.bot[37:40, 
            year := 2020]


head(ck.imts.bot)
```

Melt the table to adhere to the key-value pair schema design. 

```{r}
# Melt the data to adhere to the schema design
ck.imts.bot <- melt.data.table(ck.imts.bot,
                               id = c("year",
                                      "month"), measure = c("exports-fob-domestic",
                                                            "exports-fob-reexport", 
                                                            "exports-fob-total",
                                                            "imports-cif",
                                                            "trade-balance"))

head(ck.imts.bot)
```

Create Date from year and month. 

```{r}
# Concatenate
ck.imts.bot[, 
            date := paste0(gsub(" ", "", year, fixed = TRUE),
                           "-",
                           gsub(" ", "",month, fixed = TRUE),
                           "-",
                           "01")]


# Format Date
ck.imts.bot[, 
            date := as.Date(date, format = "%Y-%B-%d")]


# Remove year and month fields 
ck.imts.bot[, 
            year := NULL][, 
                          month := NULL]


head(ck.imts.bot)
```

Rename "variable" to "name". 

```{r}
# Rename variable to name 
setnames(ck.imts.bot, 
         "variable",
         "name")


head(ck.imts.bot)
```

Add metric key, frequency, country, category, source and property informations. 

```{r}
# Metric 
ck.imts.bot[, 
            `:=` (frequency = "monthly", 
                  country = "ck", 
                  category = "trade", 
                  source = "imts", 
                  properties = "{'currency' = 'nzd 000'}",
                  metric_key = paste0("ck","-",name,"-",date))]


head(ck.imts.bot)
```

Reorder data. 

```{r}
# Change order of columns to align with schema
ck.imts.bot <- ck.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]

head(ck.imts.bot)
```


# Read IMTS Balance of Trade Solomon Islands Data 

Read in Balance of Trade - All Items for Solomon Islands.

```{r}
# Read in Fiji File 
sl.imts.bot <- read.xlsx(xlsxFile = './data/external/Solomon_Islands_IMTS_Tables_2020.xlsx',
                         sheet = 1, 
                         startRow = 25,
                         skipEmptyCols = TRUE,
                         detectDates = TRUE,
                         rows = c(25:41),
                         cols = c(1:7))



head(sl.imts.bot)
```

Add and fix missing variable names. 

```{r}
# sl.imts.bot to data.table 
sl.imts.bot <- data.table(sl.imts.bot)


# Fix Monthly label to Year
setnames(sl.imts.bot, 
         "Monthly",
         "year")


# X2 should be Month 
setnames(sl.imts.bot, 
         "X2",
         "month")


# Exports FOB Domestic 
setnames(sl.imts.bot, 
         "X3",
         "exports-fob-domestic")


# Exports FOB Re-Export 
setnames(sl.imts.bot, 
         "X4",
         "exports-fob-reexport")


# Exports FOB Total 
setnames(sl.imts.bot, 
         "X5",
         "exports-fob-total")


# Imports CIF
setnames(sl.imts.bot, 
         "X6",
         "imports-cif")



# Trade Balance
setnames(sl.imts.bot, 
         "X7",
         "trade-balance")

head(sl.imts.bot)
```

Update Year values.

```{r}
# Update by using row number references
# 2018
sl.imts.bot[1:12, 
            year := 2018]


# 2019
sl.imts.bot[13:15, 
            year := 2019]


head(sl.imts.bot)
```

Melt the table to adhere to the key-value pair schema design. 

```{r}
# Melt the data to adhere to the schema design
sl.imts.bot <- melt.data.table(sl.imts.bot,
                               id = c("year",
                                      "month"), measure = c("exports-fob-domestic",
                                                            "exports-fob-reexport", 
                                                            "exports-fob-total",
                                                            "imports-cif",
                                                            "trade-balance"))

head(sl.imts.bot)
```

Create Date from year and month. 

```{r}
# Concatenate
sl.imts.bot[, 
            date := paste0(gsub(" ", "", year, fixed = TRUE),
                           "-",
                           gsub(" ", "",month, fixed = TRUE),
                           "-",
                           "01")]


# Format Date
sl.imts.bot[, 
            date := as.Date(date, format = "%Y-%B-%d")]


# Remove year and month fields 
sl.imts.bot[, 
            year := NULL][, 
                          month := NULL]


head(sl.imts.bot)
```

Rename "variable" to "name". 

```{r}
# Rename variable to name 
setnames(sl.imts.bot, 
         "variable",
         "name")


head(sl.imts.bot)
```

Add metric key, frequency, country, category, source and property informations. 

```{r}
# Metric 
sl.imts.bot[, 
            `:=` (frequency = "monthly", 
                  country = "sb", 
                  category = "trade", 
                  source = "imts", 
                  properties = "{'currency' = 'sbd 000'}",
                  metric_key = paste0("sb","-",name,"-",date))]


head(sl.imts.bot)
```

Reorder data. 

```{r}
# Change order of columns to align with schema
sl.imts.bot <- sl.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]


head(sl.imts.bot)
```

# Read Samoa Bureau of Statistics Trade Data 

This data has been extracted manual from the reference source outlined earlier. 

The process followed: 

1. Extract data from PDF
2. Manipulate into the required data schema using MS Excel. 
3. Save as .csv to be uploaded to PostgreSQL Database. 

```{r}
# Read in Samoa File 
ws.imts.bot <- fread('./data/interim/Samoa IMTS.csv')


head(ws.imts.bot)
```

Manipulate to fit required data schema. Add appropriate attributes and reorder the dataset to match the required schema. 

```{r}
# Keep only date, source, name, value, frequency
ws.imts.bot[, 
            `:=` (V1 = NULL,
                  country = NULL,
                  currency = NULL)]

# Metric 
ws.imts.bot[, 
            `:=` (country = "ws", 
                  properties = "{'currency' = 'wst 000'}",
                  metric_key = paste0("ws","-",name,"-",date))]


# Change order of columns to align with schema
ws.imts.bot <- ws.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]

head(ws.imts.bot)
```

# Read Vanuatu National Statistics Trade Data

This data has been extracted manual from the reference source outlined earlier. 

The process followed: 

1. Extract data from PDF (Page 6, Table 1)
2. Manipulate into the required data schema using MS Excel. 
3. Save as .csv to be uploaded to PostgreSQL Database. 

```{r}
# Read in Vanuatu File 
vn.imts.bot <- fread('./data/interim/Vanuatu IMTS cleaned.csv')


head(vn.imts.bot)
```

Manipulate to fit required data schema. Add appropriate attributes and reorder the dataset to match the required schema. 

```{r}
# Keep only date, source, name, value, frequency
vn.imts.bot[, 
            `:=` (country = NULL,
                  category = NULL, 
                  port = NULL, 
                  currency = NULL)]

# Metric 
vn.imts.bot[, 
            `:=` (country = "vn", 
                  category = "trade", 
                  properties = "{'currency' = 'million vatu'}",
                  metric_key = paste0("vn","-",name,"-",date))]


# Change order of columns to align with schema
vn.imts.bot <- vn.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]

head(vn.imts.bot)
```

Where the the values are for annual frequency, update the date field if it only contains a year value. Make formatting of the date field consistent. 

```{r}
# Update date 
vn.imts.bot[date == "2016",
            date := "2016-12-01"]

# Concatenate date with just month and year to create full date like other datasets
vn.imts.bot[frequency != "annually", 
            date := paste0("01-",
                           date)]


# Format Date frequency annual
vn.imts.bot.a <- vn.imts.bot[frequency == "annually"][, 
                                                      date := as.Date(date)]


# Format Date frequency all others
vn.imts.bot.b <- vn.imts.bot[frequency != "annually"][, 
                                                      date := as.Date(date, format = "%d-%b-%y")]


# Stack back together 
vn.imts.bot <- rbind(vn.imts.bot.a, 
                         vn.imts.bot.b)

# Remove temp ds 
rm(vn.imts.bot.a,
   vn.imts.bot.b)


head(vn.imts.bot)
```

# Read IMTS Balance of Trade Palau Data 

Read in Balance of Trade - All Items for Palau.

```{r}
# Read in Fiji File 
pl.imts.bot <- read.xlsx(xlsxFile = './data/external/Palau_IMTS_Tables_2020_.xlsx',
                         sheet = 1, 
                         startRow = 8,
                         skipEmptyCols = TRUE,
                         detectDates = TRUE,
                         rows = c(8:34),
                         cols = c(1:7))

head(pl.imts.bot)
```

Add and fix missing variable names. 

```{r}
# pl.imts.bot to data.table 
pl.imts.bot <- data.table(pl.imts.bot)


# Fix Monthly label to Year
setnames(pl.imts.bot, 
         "Monthly",
         "year")


# X2 should be Month 
setnames(pl.imts.bot, 
         "X2",
         "month")


# Exports FOB Domestic 
setnames(pl.imts.bot, 
         "X3",
         "exports-fob-domestic")


# Exports FOB Re-Export 
setnames(pl.imts.bot, 
         "X4",
         "exports-fob-reexport")


# Exports FOB Total 
setnames(pl.imts.bot, 
         "X5",
         "exports-fob-total")


# Imports CIF
setnames(pl.imts.bot, 
         "X6",
         "imports-cif")



# Trade Balance
setnames(pl.imts.bot, 
         "X7",
         "trade-balance")


head(pl.imts.bot)
```

Update Year values.

```{r}
# Update by using row number references
# 2014
pl.imts.bot[1:12, 
            year := 2018]


# 2015
pl.imts.bot[13:24, 
            year := 2019]


head(pl.imts.bot)
```

Melt the table to adhere to the key-value pair schema design. 

```{r}
# Melt the data to adhere to the schema design
pl.imts.bot <- melt.data.table(pl.imts.bot,
                               id = c("year",
                                      "month"), measure = c("exports-fob-domestic",
                                                            "exports-fob-reexport", 
                                                            "exports-fob-total",
                                                            "imports-cif",
                                                            "trade-balance"))


head(pl.imts.bot)
```

Create Date from year and month. 

```{r}
# Concatenate
pl.imts.bot[, 
            date := paste0(gsub(" ", "", year, fixed = TRUE),
                           "-",
                           gsub(" ", "",month, fixed = TRUE),
                           "-",
                           "01")]


# Format Date
pl.imts.bot[, 
            date := as.Date(date, format = "%Y-%B-%d")]


# Remove year and month fields 
pl.imts.bot[, 
            year := NULL][, 
                          month := NULL]


head(pl.imts.bot)
```

Rename "variable" to "name". 

```{r}
# Rename variable to name 
setnames(pl.imts.bot, 
         "variable",
         "name")


head(pl.imts.bot)
```

Add metric key, frequency, country, category, source and property informations. 

```{r}
# Metric 
pl.imts.bot[, 
            `:=` (frequency = "monthly", 
                  country = "pl", 
                  category = "trade", 
                  source = "imts", 
                  properties = "{'currency' = 'usd'}",
                  metric_key = paste0("pl","-",name,"-",date))]


head(pl.imts.bot)
```

Reorder data. 

```{r}
# Change order of columns to align with schema
pl.imts.bot <- pl.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]


head(pl.imts.bot)
```

# Read IMTS Balance of Trade Kiribati Data

Read in Balance of Trade - All Items for Kiribati

```{r}
# Read in Kiribati File 
ki.imts.bot <- read.xlsx(xlsxFile = './data/external/Kiribati_IMTS_Tables_2020.xlsx',
                         sheet = 1, 
                         startRow = 14,
                         skipEmptyCols = TRUE,
                         detectDates = TRUE,
                         rows = c(14:67),
                         cols = c(1:7))


head(ki.imts.bot)
```

Add and fix missing variable names. 

```{r}
# ki.imts.bot to data.table 
ki.imts.bot <- data.table(ki.imts.bot)


# Remove empty record
ki.imts.bot <- ki.imts.bot[!is.na(X2)]



# Fix Monthly label to Year
setnames(ki.imts.bot, 
         "Monthly",
         "year")


# X2 should be Month 
setnames(ki.imts.bot, 
         "X2",
         "month")


# Exports FOB Domestic 
setnames(ki.imts.bot, 
         "X3",
         "exports-fob-domestic")


# Exports FOB Re-Export 
setnames(ki.imts.bot, 
         "X4",
         "exports-fob-reexport")


# Exports FOB Total 
setnames(ki.imts.bot, 
         "X5",
         "exports-fob-total")


# Imports CIF
setnames(ki.imts.bot, 
         "X6",
         "imports-cif")



# Trade Balance
setnames(ki.imts.bot, 
         "X7",
         "trade-balance")


head(ki.imts.bot)
```

Update Year values.

```{r}
# Update by using row number references
# 2014
ki.imts.bot[1:12, 
            year := 2016]


# 2015
ki.imts.bot[13:24, 
            year := 2017]


# 2016
ki.imts.bot[25:36, 
            year := 2018]


# 2017
ki.imts.bot[37:48, 
            year := 2019]


head(C)
```

Melt the table to adhere to the key-value pair schema design. 

```{r}
# Melt the data to adhere to the schema design
ki.imts.bot <- melt.data.table(ki.imts.bot,
                               id = c("year",
                                      "month"), measure = c("exports-fob-domestic",
                                                            "exports-fob-reexport", 
                                                            "exports-fob-total",
                                                            "imports-cif",
                                                            "trade-balance"))


head(ki.imts.bot)
```

Create Date from year and month. 

```{r}
# Concatenate
ki.imts.bot[, 
            date := paste0(gsub(" ", "", year, fixed = TRUE),
                           "-",
                           gsub(" ", "",month, fixed = TRUE),
                           "-",
                           "01")]


# Format Date
ki.imts.bot[, 
            date := as.Date(date, format = "%Y-%B-%d")]


# Remove year and month fields 
ki.imts.bot[, 
            year := NULL][, 
                          month := NULL]


head(ki.imts.bot)
```

Rename "variable" to "name". 

```{r}
# Rename variable to name 
setnames(ki.imts.bot, 
         "variable",
         "name")


head(ki.imts.bot)
```

Add metric key, frequency, country, category, source and property informations. 

```{r}
# Metric 
ki.imts.bot[, 
            `:=` (frequency = "monthly", 
                  country = "ki", 
                  category = "trade", 
                  source = "imts", 
                  properties = "{'currency' = 'aud 000'}",
                  metric_key = paste0("ki","-",name,"-",date))]


head(ki.imts.bot)
```

Reorder data. 

```{r}
# Change order of columns to align with schema
ki.imts.bot <- ki.imts.bot[,
                           .(metric_key,
                             frequency, 
                             country,
                             category, 
                             source,
                             name,
                             date, 
                             value,
                             properties)]


head(ki.imts.bot)
```



# Append to PostgresSQL Database 

Connect to the hosted PostgreSQL database. 

Table used is country_metrics. 

## Fiji IMTS Balance of Trade Upload

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc1 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs1 <- dbFetch(rc1)



# # Append fj.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         fj.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs2 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs2 <- dbFetch(rs2)


dbDisconnect(db.con)
```

## Solomon Islands IMTS Balance of Trade Upload

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc3 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs3 <- dbFetch(rc3)


# # Append sl.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         sl.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs4 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs4 <- dbFetch(rs4)


dbDisconnect(db.con)
```

## Cook Islands IMTS Balance of Trade Upload

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc5 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs5 <- dbFetch(rc5)


# # Append ck.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         ck.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs6 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs6 <- dbFetch(rs6)


dbDisconnect(db.con)
```

## Vanuatu IMTS VNS Balance of Trade Upload

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc7 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs7 <- dbFetch(rc7)


# # Append vn.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         vn.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)
# 

# Count rows after upload
rs8 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs8 <- dbFetch(rs8)


dbDisconnect(db.con)
```

## Samoa IMTS SBS Balance of Trade Upload 

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc9 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs9 <- dbFetch(rc9)


# # Append ws.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         ws.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs10 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs10 <- dbFetch(rs10)


dbDisconnect(db.con)

```

## Palau IMTS Balance of Trade Upload 

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc11 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs11 <- dbFetch(rc11)


# # Append pl.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         pl.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs12 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs12 <- dbFetch(rs12)


dbDisconnect(db.con)

```

## Kiribati IMTS Balance of Trade Upload 

```{r eval=FALSE}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
rc13 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs13 <- dbFetch(rc13)


# # Append ki.imts.bot to country_metrics
# RPostgres::dbWriteTable(db.con,
#                         "country_metrics",
#                         ki.imts.bot,
#                         append = TRUE,
#                         row.names = FALSE)


# Count rows after upload
rs14 <- RPostgres::dbSendQuery(db.con, "SELECT count(*) as count FROM country_metrics")


count_recs14 <- dbFetch(rs14)


dbDisconnect(db.con)

```

## Clean up 

Remove temp variables used to test.

```{r eval=FALSE}
# Remove vars not needed.
rm(rc1,
   rs2, 
   rc3,
   rs4,
   rc5,
   rs6,
   rc7,
   rs8,
   rc9,
   rs10)
```

# Final Result: country_metrics

```{r}
# Connect to a specific postgres database
db.con <- dbConnect(RPostgres::Postgres(),
                    dbname = 'aishackathon', 
                    host = 'ais-hack.cirquhp75zcc.us-east-2.rds.amazonaws.com', 
                    port = 5432, 
                    user = Sys.getenv("userid"),
                    password = Sys.getenv("pwd"))


# Count rows before upload
country_metrics <- data.table(RPostgres::dbGetQuery(db.con, "SELECT * FROM country_metrics"))

stg_trade_agg <- data.table(RPostgres::dbGetQuery(db.con, "SELECT * FROM stg_trade_agg"))

dbListTables(db.con)

View(country_metrics[country == "ck" & category == "tourism", 
                     .(s)])


# Preview
tail(country_metrics)
```
